---
title: "MEDS Alumni Analysis"
author: "Sofia Sarak"
format: html
---

The following document analyzes the careers of MEDS alumni after graduation. The following explorations were conducted:

- the number of roles an individual has held since graduating (total as well as per year)
- how many of the Class of 2022 are still at their first placement
- average length of job search (as well as by cohort)
- mobility between sectors
- differences in current placement sector, by cohort
- % of alumni with missing data
- common job titles: analyst, scientist, specialist, planner, engineer, manager

```{r}
#| message: false
#| warning: false
#| echo: false

# save Bren colors
blue <- "#14375C"
green <- "#61A845"

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      data import and initial wrangling                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggsankey)

# read in data
data <- read_csv(here("data", "meds_2020-26_placement_data.csv")) %>% 
  clean_names()

# encode "N/A" as actual NA values

data <- data %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "Unknown")))

```

## Placements Since Graduation

This analysis considers promotions within the same organization as a placement.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##        placement data wrangling - create binary placement variable       ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# create placement data frame (for binary vars)
placement <- data %>% 
  
  # keep only necessary columns (a placement job title signifies placement)
  select(year, last_name, first_name, 
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 

# keep only identifiers and binary vars
  
  select(year, last_name, first_name, initial_bin, second_bin, third_bin, fourth_bin)

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                find total number of placements per individual            ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement <- placement %>% 
  
  # sum columns for total positions
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                  find placements per individual per year                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement <- placement %>% 
  
  mutate(
    
    # calculate number of years since grad (years in workforce)
    workforce_years = 2026 - year,
         
    # divide placements by number of years in workforce
    placements_per_year = case_when(
      
      # for MEDS class of 2026, default to total placement (otherwise, 1/0 = Inf)
      workforce_years == 0 ~ total_placements,
      
      # division for other years
      .default = total_placements/workforce_years))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placements_plot <- placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(total_placements)) +
    geom_bar(fill = blue,) +
  labs(x = "Number of Placements Since Graduation",
       y = "Number of Graduates",
       title = "Job Mobility in MEDS Graduates") +
  theme_minimal()

yearly_placements_plot <- placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(placements_per_year)) +
    geom_bar(fill = blue) +
  labs(x = "Number of Placements per Year",
       y = "Number of Graduates",
       title = "Job Mobility in MEDS Graduates",
       subtitle = "(Yearly Average)") +
  theme_minimal()

placements_plot + yearly_placements_plot
```
The majority of graduates switch positions once a year or less. This is skewed heavily by graduates who have only had a year or two in the workforce.

## Class of 2022: Have they retained their first placement?

Here, we isolate the Class of 2022 -- they are the first MEDS cohort, and have had 5 years in the workforce. This provides a more standardized look at job mobility over time.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##            create a placement data frame just for class of 2022          ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# this version of the df filters for 2022 and also keeps organization name variables

first_cohort <- data %>% 
  
  # filter for class of 2022
  filter(year == 2022) %>% 
  
  # select necessary columns
  select(year, last_name, first_name, initial_placement_organization_name, current_most_recent_position_organization_name,
         
         # placement columns to know if there were any promotions
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values SOMETHING WRONG WITH IT THIS TIME?
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 
  
  select(year, last_name, first_name,  initial_placement_organization_name, current_most_recent_position_organization_name,
         initial_bin, second_bin, third_bin, fourth_bin) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                  find individuals at their original orgs                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

original_org <- first_cohort %>% 
  
  # keep only obvs where initial and current org are the same
  filter(initial_placement_organization_name == current_most_recent_position_organization_name) %>% 
  
  # add column for the total number of promotions/position changes
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))

# add columnn to first_cohort data frame that denotes whether individual is still at their original org

first_cohort <- first_cohort %>% 
  
  # create a binary column, =1 when original org is equal to current org
  mutate(original_org = case_when(
    initial_placement_organization_name == current_most_recent_position_organization_name ~ 1,
    .default = 0))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# bar plot comparing original org to not original org
original_org_plot <- ggplot(first_cohort, aes(x = as.factor(original_org))) +
  geom_bar(fill = blue) +
  
  # add labels for each bar
  geom_text(stat = "count", aes(label = after_stat(count), y = after_stat(count)),
            vjust = 2, size = 5, color = "white") +
  labs(y = "Number of Graduates",
       x = " ",
       title = "How many of MEDS Class of 2022 are still at their initial placements?",
       subtitle = "Of the 7 still at their initial placement, 4 still hold the same job title.") +
  scale_x_discrete(labels = c("1" = "Initial placement", "0" = "Not at initial placement")) +
  theme_minimal()

# number of placements in the class of 2022
first_cohort_placements_plot <- placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # filter for class of 2022
  filter(year == 2022) %>% 
  
  # plot distribution
  ggplot(aes(total_placements)) +
    geom_bar(fill = blue,) +
   # add labels for each bar
  geom_text(stat = "count", aes(label = after_stat(count), y = after_stat(count)),
            vjust = 2, size = 5, color = "white") +
  labs(x = "Number of Placements Since Graduation",
       y = "Number of Graduates",
       title = "Job Mobility in 2022 MEDS Graduates") +
  theme_minimal()

original_org_plot / first_cohort_placements_plot

```

The majority of the class of 2022 is not still at their initial organization, and have had 2 placements since graduating.

## Length of Job Search

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        add job search length columns                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

search <- data %>% 
  
  # select necessary columns
  select(year, last_name, first_name, ends_with("date")) %>% 
  
  # transform date columns to date format
  mutate(across(ends_with("date"), my)) %>% 
  
  # add column for graduation day approximation (06 + `year` + 15)
  mutate(grad_date = paste0("06/", year, "/15"),
         grad_date = myd(grad_date)) %>% 
  
  # calculate amount of time between graduation and initial placement
  mutate(initial_search_days = as.numeric(initial_placement_start_date - grad_date),
         
         # fix so that if job was started in July (14 days or earlier), search length is actually 0
         initial_search_days = if_else(initial_search_days < 15, 0, initial_search_days))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##             filter for jobs found before/after grad, or at all           ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Create binary column where positions found after graduation are denoted with 1
search <- search %>%
  
  # add column
  mutate(after_grad = case_when(
    initial_search_days > 0 ~ 1,
    .default = 0))

# For a more accurate representation of job search length, we only select for those individuals that have found positions/the positions are publicly available
search_non_na <- search %>% 
  filter(!is.na(initial_search_days))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# job search distributions
job_search_dist_plot <- search_non_na %>% 
  ggplot(aes(x = initial_search_days)) + 
  geom_histogram(fill = blue) +
  labs(x = "Initial Job Search Length (Days)",
       y = "Number of Graduates",
       title = "Distribution of Job Search Lengths in MEDS Graduates",
       subtitle = "Average job search length of 62 days is shown in red; before graduation finds excluded from plot") + 
  theme_minimal() +
  
  # add line for average
  geom_vline(xintercept = mean(search$initial_search_days, na.rm = TRUE),
             color = "red",
             linetype = "dashed")

# number of roles found before and after graduation
after_grad_plot <- ggplot(search_non_na, aes(x = as.factor(after_grad))) +
  geom_bar(fill = blue) +
  
  # add labels for each bar
  geom_text(stat = "count", aes(label = after_stat(count), y = after_stat(count)),
            vjust = 2, size = 5, color = "white") +
  labs(y = "Number of Graduates",
       x = " ",
       title = "When do MEDS students find their initial placements?") +
  scale_x_discrete(labels = c("1" = "After graduation", "0" = "Before graduation")) +
  theme_minimal()

job_search_dist_plot / after_grad_plot
```
These analyses rest on the assumption that any roles started in or before July were found before graduation. This can obviously introduce some error - an entire month *can* be enough time to find and begin a role, but we do not have the granularity in our data to determine when exactly in July the roles were started. 

Overall, the job search lasted around two months for most graduates, and the majority of graduates began their roles after July (or over a month after graduation).

These conclusions do not take into consideration the fact that there were individuals who had graduated but were not actively searching for roles.

## Job search by cohort

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                 find average job search by cohort and plot               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
search_non_na %>% 
  group_by(year) %>% 
  summarise(average_search = round(mean(initial_search_days, na.rm = TRUE))) %>% 
  
  ggplot(aes(x = year, y = average_search)) +
  geom_col(fill = blue) +
  geom_text(aes(label = average_search),
            color = "white",
            vjust = 1.4,
            size = 4) +
  labs(title = "Length of job search by MEDS cohort",
       x = "",
       y = "Days") +
  theme_minimal()
```
These resulls are heavily skewed by 2025 and 2026 graduates who have not find roles yet.

## Mobility between sectors

Here, I explore whether graduates switch between sectors as they change roles.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      read in and wrangle sector data                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

sector <- read_csv(here("data", "sectors.csv")) %>% 
  
  clean_names() %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "Unknown"))) %>% 
  
  # remove rows where there was no initial placement/no data (mostly class of 2026)
  filter(!is.na(initial_placement))

# convert data to ggsankey format
sector_sankey <- sector %>% 
  
  # function comes from ggsankey package
  make_long(initial_sector, second_sector, third_sector, fourth_sector) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        plot using ggsankey package                       ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

ggplot(sector_sankey, aes(x = x, # set aesthetics to make make_long() formatting
                         next_x = next_x,
                         node = node,
                         next_node = next_node,
                         fill = factor(node),
                         label = node)) +
         
         # customize 
         geom_sankey(flow.alpha = 0.5,  # opacity
                     node.color = "NA", # border color
                     show.legend = FALSE) + 
         
         # add node labels
         geom_sankey_label(size = 3, color = "black", fill= "white") +
  
        # set x-axis labels
         scale_x_discrete(labels = c("Initial\nPlacement", "Second\nPlacement", "Third\nPlacement", "Fourth\nPlacement")) +
  
        theme_bw()+
        theme(legend.position = "none",
              axis.title = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              panel.grid = element_blank()) +
  
        scale_fill_viridis_d(option = "rocket") +
  
        labs(title = "Movement across sectors, MEDS graduates",
             fill = "Nodes")
```

This diagram shows that many alumni start out in Research/Education and move to other sectors. Many who start in the corporate sector remain there.

## Missing data

Other than for initial placements, the MEDS career team does not have a way to gather data on alumni careers other than LinkedIn or other social pages, such as personal websites or GitHub bios. Unfortunately, is not uncommon for an individual's online presence to be outdated, resulting in missing data for analyses such as this one. Because of this, I briefly explore how many alumni have outdated or missing information regarding their career.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                create binary variable for unknown placements             ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# `current_position_unknown` column is NA if position IS known, and "Unknown" otherwise
data <- data %>% 
  
  # change contents of column
  mutate(current_position_unknown = case_when(
    current_position_unknown == "Unknown" ~ 1,
    .default = 0
  ))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                            plot unknown vs known                         ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~


unknown_plot <- data %>% 
  ggplot(aes(y = "", fill = as.factor(current_position_unknown))) + 
  geom_bar(position = "fill") + 
  geom_text(
    stat = "count",
    
    # calculate proportions dynamically and format them as percentages
    aes(
      label = scales::percent(after_stat(count / sum(count)), accuracy = 1),
      group = current_position_unknown
    ),
    position = position_fill(vjust = 0.5),
    size = 5,
    color = "white"
  ) +
  labs(
    y = "", 
    x = NULL, 
    title = "What percent of alumni careers do we\nactually know about?",
    fill = " "
  ) + 
  scale_fill_discrete(labels = c("1" = "Unknown", "0" = "Known"),
                      palette = c(blue, green)) +
  theme_minimal() +
  theme(axis.text = element_blank(),
        legend.position = "bottom", 
        legend.direction = "horizontal")

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                       Proportion known by class year                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

unknown_by_class_plot <- data %>% 
  
  # calculate % unknown for each year
  group_by(year) %>% 
  summarize(perc_unknown = mean(current_position_unknown, na.rm = TRUE)) %>% 
  
  # plot
  ggplot(aes(x = year, y = perc_unknown)) +
  geom_col(fill = blue) +
   geom_text(
    aes(label = paste0(round(perc_unknown * 100), "%")),
    vjust = 1.5,
    size = 5,
    color = "white"
  ) +
  theme_minimal() +
  labs(title = "Perentage Missing Data, by Cohort",
       x = "",
       y = "% Unknown")
  

unknown_plot + unknown_by_class_plot
```


With this insight, it becomes clear that the previous conclusions made are based only on a proportion of total MEDS outcomes. However, it is important to note that the majority of unknown careers come from the two most recent classes (2025 and 2026). This may be due to the fact that many may have not secured full-time roles yet.

## Position Titles

Common job titles across alumni include the words: analyst, scientist, specialist, planner, engineer, and manager, highlighting the range of roles that the MEDS program prepares graduates for.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                    create df just with position titles                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement_titles <- data %>% 
  select(contains("job_title")) %>% 
  
  # create a long data frame
  pivot_longer(cols = everything(), names_to = "placement_order", values_to = "title") %>% 
  drop_na()

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                    create binary columns noting titles                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement_titles <- placement_titles %>% 
  
 # Analyst
   mutate(analyst = case_when(
    str_detect(title, "Analyst") ~ 1,
    .default = 0)) %>% 
  
  # Scientist
   mutate(scientist = case_when(
    str_detect(title, "Scientist") ~ 1,
    .default = 0)) %>% 
  
  # Specialist
   mutate(specialist = case_when(
    str_detect(title, "Specialist") ~ 1,
    .default = 0)) %>% 
  
  # Planner
   mutate(planner = case_when(
    str_detect(title, "Planner") ~ 1,
    .default = 0)) %>% 
  
  # Engineer
   mutate(engineer = case_when(
    str_detect(title, "Engineer") ~ 1,
    .default = 0)) %>% 
  
  # Manager
   mutate(manager = case_when(
    str_detect(title, "Manager") ~ 1,
    .default = 0))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                calculate number of roles with those titles               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

titles_count <- placement_titles %>% 
  
  # sum all columns
    summarise(across(c(analyst:manager), sum)) %>% 
    pivot_longer(analyst:manager, names_to = "title", values_to = "count")

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

titles_count %>% 
  ggplot(aes(x = reorder(title, -count), y = count)) +
  geom_col(fill = blue) +
  geom_text(aes(label = count),
            color = "white",
            vjust = 1.4,
            size = 4) +
  labs(title = "Most common job titles in MEDS alumni",
       x = "",
       y = "Number of roles") +
  theme_minimal()
```
Scientists and analysts takes the lead as the most common job titles.
