---
title: "MEDS Alumni Analysis"
author: "Sofia Sarak"
format: pdf
---

This document analyzes the careers of MEDS alumni after graduation. The following explorations were conducted:

- the number of roles an individual has held since graduating (total as well as per year)
- how many of the Class of 2022 are still at their first placement
- average length of job search (as well as by cohort)
- mobility between sectors
- differences in current placement sector, by cohort
- % of alumni with missing data
- common job titles: analyst, scientist, specialist, planner, engineer, manager

```{r}
#| message: false
#| warning: false
#| echo: false

# save Bren colors
blue <- "#14375C"
green <- "#61A845"

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      data import and initial wrangling                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggtext)

# read in data
data <- read_csv(here("data", "meds_2020-26_placement_data.csv")) %>% 
  clean_names()

# encode "N/A" as actual NA values

data <- data %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "Unknown")))

```

## Length of Job Search

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        add job search length columns                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

search <- data %>% 
  
  # select necessary columns
  select(year, last_name, first_name, ends_with("date")) %>% 
  
  # transform date columns to date format
  mutate(across(ends_with("date"), my)) %>% 
  
  # add column for graduation day approximation (06 + `year` + 15)
  mutate(grad_date = paste0("06/", year, "/15"),
         grad_date = myd(grad_date)) %>% 
  
  # calculate amount of time between graduation and initial placement
  mutate(initial_search_days = as.numeric(initial_placement_start_date - grad_date),
         
         # fix so that if job was started in July (14 days or earlier), search length is actually 0
         initial_search_days = if_else(initial_search_days < 15, 0, initial_search_days))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##             filter for jobs found before/after grad, or at all           ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Create binary column where positions found after graduation are denoted with 1
search <- search %>%
  
  # add column
  mutate(after_grad = case_when(
    initial_search_days > 0 ~ 1,
    .default = 0))

# For a more accurate representation of job search length, we only select for those individuals that have found positions/the positions are publicly available
search_non_na <- search %>% 
  filter(!is.na(initial_search_days)) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# job search distributions
search_non_na %>% 
  
  filter(initial_search_days < 300) %>% 
  
  ggplot(aes(x = initial_search_days)) + 
  geom_density(fill = blue,
               color = blue,
               alpha = 0.75) +
  labs(x = "Initial Job Search Length (Days)",
       y = "Number of Graduates",
       title = "Distribution of Job Search Lengths in MEDS Graduates",
       subtitle = "The average job search length is <span style='color: Red;'>**62 days**</span>.*",
       caption = "*Roles found before graduation are excluded from plot.") + 
  theme_minimal() +
  
  theme(plot.subtitle = element_markdown()) +
  
  # add line for average
  geom_vline(xintercept = mean(search$initial_search_days, na.rm = TRUE),
             color = "red",
             linetype = "dashed")

# check y axis?? also think about adding july job starts back in
```
These analyses rest on the assumption that any roles started in or before July were found before graduation. This can obviously introduce some error - an entire month *can* be enough time to find and begin a role, but we do not have the granularity in our data to determine when exactly in July the roles were started. 

Overall, the job search lasted around two months for most graduates, and the majority of graduates began their roles after July (or over a month after graduation).

These conclusions do not take into consideration the fact that there were individuals who had graduated but were not actively searching for roles.



```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      read in and wrangle sector data                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

sector <- read_csv(here("data", "sectors.csv")) %>% 
  
  clean_names() %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "Unknown"))) %>% 
  
  # remove rows where there was no initial placement/no data (mostly class of 2026)
  filter(!is.na(initial_placement))

# convert data to ggsankey format
sector_sankey <- sector %>% 
  
  # function comes from ggsankey package
  make_long(initial_sector, second_sector, third_sector, fourth_sector) 

```

This diagram shows that many alumni start out in Research/Education and move to other sectors. Many who start in the corporate sector remain there.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##          attach sector to complete df for sector by year analysis        ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# separate sector df for joining 
initial <- sector %>% 
  select(initial_placement, initial_sector)
second <- sector %>% 
  select(second_placement, second_sector)
third <- sector %>% 
  select(third_placement, third_sector)
fourth <- sector %>% 
  select(fourth_placement, fourth_sector)

# join separate dfs to data (takes a sec)
data <- data %>% 
  left_join(initial, by=c("initial_placement_organization_name" = "initial_placement")) %>% 
  distinct()
data <- data %>% 
  left_join(second, by=c("second_placement_organization_name" = "second_placement")) %>% 
  distinct()
data <- data %>% 
  left_join(third, by=c("third_placement_organization_name" = "third_placement")) %>% 
  distinct()
data <- data %>% 
  left_join(fourth, by=c("fourth_placement_organization_name" = "fourth_placement")) %>% 
  distinct()

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                            group by class year                           ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
sector_by_cohort <- data %>% 
  pivot_longer(cols = c(initial_sector:fourth_sector),
               values_to = "sector",
               names_to = "placement_order") %>% 
  group_by(year, sector) %>% 
  summarize(count = n()) %>% 
  filter(!is.na(sector))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    plot                                  ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# faceted by year
# ggplot(sector_by_cohort, aes(y = sector, x = count)) +
#   geom_col() +
#   facet_wrap(~year)

ggplot(sector_by_cohort, aes(y = as.factor(year), x = count, fill = sector)) +
  geom_col() +
  labs(x = "Number of positions",
       y = "",
       fill = "Sector", 
       title = "Differences in sector by cohort") +
  theme_minimal()

# could also benefit with plotting by percentage?
```


## Position Titles

Common job titles across alumni include the words: analyst, scientist, specialist, planner, engineer, and manager, highlighting the range of roles that the MEDS program prepares graduates for.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                    create df just with position titles                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement_titles <- data %>% 
  select(contains("job_title")) %>% 
  
  # create a long data frame
  pivot_longer(cols = everything(), names_to = "placement_order", values_to = "title") %>% 
  drop_na()

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                    create binary columns noting titles                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement_titles <- placement_titles %>% 
  
 # Analyst
   mutate(analyst = case_when(
    str_detect(title, "Analyst") ~ 1,
    .default = 0)) %>% 
  
  # Scientist
   mutate(scientist = case_when(
    str_detect(title, "Scientist") ~ 1,
    .default = 0)) %>% 
  
  # Specialist
   mutate(specialist = case_when(
    str_detect(title, "Specialist") ~ 1,
    .default = 0)) %>% 
  
  # Planner
   mutate(planner = case_when(
    str_detect(title, "Planner") ~ 1,
    .default = 0)) %>% 
  
  # Engineer
   mutate(engineer = case_when(
    str_detect(title, "Engineer") ~ 1,
    .default = 0)) %>% 
  
  # Manager
   mutate(manager = case_when(
    str_detect(title, "Manager") ~ 1,
    .default = 0))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                calculate number of roles with those titles               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

titles_count <- placement_titles %>% 
  
  # sum all columns
    summarise(across(c(analyst:manager), sum)) %>% 
    pivot_longer(analyst:manager, names_to = "title", values_to = "count")

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

titles_count %>% 
  ggplot(aes(x = reorder(title, -count), y = count)) +
  geom_col(fill = blue) +
  geom_text(aes(label = count),
            color = "white",
            vjust = 1.4,
            size = 4) +
  labs(title = "Most common job titles in MEDS alumni",
       x = "",
       y = "Number of roles") +
  theme_minimal()
```
Scientists and analysts takes the lead as the most common job titles.
