---
title: "Job Search Length Analysis"
author: "Sofia Sarak"
format: html
editor: visual
---

```{r}
#| message: false
#| warning: false
# load necessary libraries
library(tidyverse)
library(here)
library(janitor)
```


```{r}
#| message: false
# read in data
data <- read_csv(here("data", "meds_2020-26_placement_data.csv")) %>% 
  clean_names()
```

## Number of Placements

### Data Wrangling

Encode "N/A" as actual NA values.

```{r}
data <- data %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "Unknown")))
```

Create binary variable to describe job placement. If an individual has a fourth placement, the value for that column is 1. If an individual only has an initial placement, the initial column is 1, and the rest (second, third, and fourth placement) are 0. In this way, summing across binary columns can quickly reveal the number of positions held.

```{r}
# create placement data frame (for binary vars)
placement <- data %>% 
  
  # keep only necessary columns (a placement job title signifies placement)
  select(year, last_name, first_name, 
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 

# keep only identifiers and binary vars
  
  select(year, last_name, first_name, initial_bin, second_bin, third_bin, fourth_bin)
```

Find the total number of positions per individual (by summing binary variables). This can offer insights on career progression within an organization, environmental data science market dynamics, or even job satisfaction.

```{r}
placement <- placement %>% 
  
  # sum columns for total positions
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))
```

This value is of course very dependent on the number of years since the individual's graduation. We can find an average number of jobs per year value by 1) Calculating years since graduation (2026 - `year`), and 2) Dividing `total_placements` by `year`.

```{r}
placement <- placement %>% 
  
  mutate(
    
    # calculate number of years since grad (years in workforce)
    workforce_years = 2026 - year,
         
    # divide placements by number of years in workforce
    placements_per_year = case_when(
      
      # for MEDS class of 2026, default to total placement (otherwise, 1/0 = Inf)
      workforce_years == 0 ~ total_placements,
      
      # division for other years
      .default = total_placements/workforce_years))
```


### Visualization

```{r}
# save Bren blue
blue <- "#14375C"
```


What is the distribution of career swicthes among graduates?
```{r}
placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(placements_per_year)) +
    geom_histogram(fill = blue) +
  labs(x = "Number of Placements per Year",
       y = "Frequency",
       title = "Job Movement (?) in MEDS Graduates") +
  theme_minimal()
```
The majority of graduates switch positions once a year or less.

## Length of Job Search

### Data Wrangling

In order to estimate the length of each graduates job search, we use just the `start_date` and `end_date` columns for each placement. We can then add a graduation column (using June 15th as a default graduation date), and calculating the difference between initial placement start date and graduation.

These results may not be completely accurate, as a jobs could have been secured weeks, if not months, before graduation.

```{r}
search <- data %>% 
  
  # select necessary columns
  select(year, last_name, first_name, ends_with("date")) %>% 
  
  # transform date columns to date format
  mutate(across(ends_with("date"), my)) %>% 
  
  # add column for graduation day approximation (06 + `year` + 15)
  mutate(grad_date = paste0("06/", year, "/15"),
         grad_date = myd(grad_date)) %>% 
  
  # calculate amount of time between graduation and initial placement
  mutate(initial_search = case_when(
    initial_placement_start_date - grad_date == "-14 days" ~ 0,
    .default = initial_placement_start_date - grad_date))
```

