---
title: "Job Search Length Analysis"
author: "Sofia Sarak"
format: html
editor: visual
---

The following document utilizes the MEDS alumni database to perform analyses exploring job mobility and length of job search.

## Setup

```{r}
#| message: false
#| warning: false
# load necessary libraries
library(tidyverse)
library(here)
library(janitor)
```

```{r}
#| message: false
# read in data
data <- read_csv(here("data", "meds_2020-26_placement_data.csv")) %>% 
  clean_names()
```

Encode "N/A" as actual NA values.

```{r}
data <- data %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "Unknown")))
```

## Number of Placements

### Data Wrangling

Create binary variable to describe job placement. If an individual has a fourth placement, the value for that column is 1. If an individual only has an initial placement, the initial column is 1, and the rest (second, third, and fourth placement) are 0. In this way, summing across binary columns can quickly reveal the number of positions held.

```{r}
# create placement data frame (for binary vars)
placement <- data %>% 
  
  # keep only necessary columns (a placement job title signifies placement)
  select(year, last_name, first_name, 
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 

# keep only identifiers and binary vars
  
  select(year, last_name, first_name, initial_bin, second_bin, third_bin, fourth_bin)
```

Find the total number of positions per individual (by summing binary variables). This can offer insights on career progression within an organization, environmental data science market dynamics, or even job satisfaction.

```{r}
placement <- placement %>% 
  
  # sum columns for total positions
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))
```

This value is of course very dependent on the number of years since the individual's graduation. We can find an average number of jobs per year value by 1) Calculating years since graduation (2026 - `year`), and 2) Dividing `total_placements` by `year`.

```{r}
placement <- placement %>% 
  
  mutate(
    
    # calculate number of years since grad (years in workforce)
    workforce_years = 2026 - year,
         
    # divide placements by number of years in workforce
    placements_per_year = case_when(
      
      # for MEDS class of 2026, default to total placement (otherwise, 1/0 = Inf)
      workforce_years == 0 ~ total_placements,
      
      # division for other years
      .default = total_placements/workforce_years))
```


### Visualization

```{r}
# save Bren blue
blue <- "#14375C"
```


What is the distribution of career switches among graduates?
```{r}
placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(placements_per_year)) +
    geom_histogram(fill = blue) +
  labs(x = "Number of Placements per Year",
       y = "Frequency",
       title = "Job Mobility in MEDS Graduates") +
  theme_minimal()
```
The majority of graduates switch positions once a year or less.

### Class of 2022

Here, we isolate the Class of 2022 -- they are the first MEDS cohort, and have had 5 years in the workforce. We repeat the wrangling that is under the *Number of Placements* analysis, where we create a binary variable to find the number of position held.

```{r}
first_class <- data %>% 
  
  # filter for class of 2022
  filter(year == 2022) %>% 
  
  # select necessary columns
  select(year, last_name, first_name, initial_placement_organization_name, current_most_recent_position_organization_name,
         
         # placement columns to know if there were any promotions
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values SOMETHING WRONG WITH IT THIS TIME?
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 
  
  select(year, last_name, first_name,  initial_placement_organization_name, current_most_recent_position_organization_name,
         initial_bin, second_bin, third_bin, fourth_bin) 
```

Next, we explore how many MEDS graduates stay with the employer they had their initial placement with. A separate dataframe is created.

```{r}
# find individuals who are still at their original org
original_org <- first_class %>% 
  
  # keep only obvs where initial and current org are the same
  filter(initial_placement_organization_name == current_most_recent_position_organization_name) %>% 
  
  # add column for the total number of promotions/position changes
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))
```

In order to understand our findings, proportions are presented in print statements.

```{r}
# find number of individuals at original org, versus total number
print(paste(nrow(original_org), "of the original", nrow(first_class), "MEDS graduates are still at their first placement."))
print(paste("That's", nrow(original_org)/nrow(first_class) * 100, "% of MEDS 2022 graduates."))

original_org_same_title <- original_org %>% 
  filter(total_placements == 1)

print(paste("Of this proportion,", nrow(original_org_same_title), "still hold the same job title."))
```

### Apply to all years (for fun)

The Class of 2022 offers the most time in the workforce post-graduation. However, we can apply the analysis to all MEDS cohorts. Interpretations of these results should include the caveat that time plays a large factor in whether an individual is still with their original employer or not.

```{r}
placement_1 <- data %>% 
  
  select(year, last_name, first_name, initial_placement_organization_name, current_most_recent_position_organization_name,
         
         # placement columns to know if there were any promotions
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 

# keep only identifiers and binary vars
  
   select(year, last_name, first_name,  initial_placement_organization_name, current_most_recent_position_organization_name,
         initial_bin, second_bin, third_bin, fourth_bin) 

# find individuals who are still at their original org
original_org_all <- placement_1 %>% 
  
  # keep only obvs where initial and current org are the same
  filter(initial_placement_organization_name == current_most_recent_position_organization_name) %>% 
  
  # add column for the total number of promotions/position changes
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))

# find number of individuals at original org, versus total number
print(paste(nrow(original_org_all), "of the original", nrow(placement_1), "MEDS graduates are still at their first placement."))
print(paste("That's", round(nrow(original_org_all)/nrow(placement_1) * 100), "% of all MEDS graduates."))

original_org_all_same_title <- original_org_all %>% 
  filter(total_placements == 1)

print(paste("Of this proportion,", nrow(original_org_all_same_title), "still hold the same job title."))
```

## Length of Job Search

### Data Wrangling

In order to estimate the length of each graduates job search, we use just the `start_date` and `end_date` columns for each placement. We can then add a graduation column (using June 15th as a default graduation date), and calculating the difference between initial placement start date and graduation.

These results may not be completely accurate, as a jobs could have been secured weeks, if not months, before graduation.

```{r}
search <- data %>% 
  
  # select necessary columns
  select(year, last_name, first_name, ends_with("date")) %>% 
  
  # transform date columns to date format
  mutate(across(ends_with("date"), my)) %>% 
  
  # add column for graduation day approximation (06 + `year` + 15)
  mutate(grad_date = paste0("06/", year, "/15"),
         grad_date = myd(grad_date)) %>% 
  
  # calculate amount of time between graduation and initial placement
  mutate(initial_search_days = as.numeric(initial_placement_start_date - grad_date),
         
         # fix so that if job was started in July (14 days or earlier), search length is actually 0
         initial_search_days = if_else(initial_search_days < 15, 0, initial_search_days))
```

### Summary Stats and Viz

```{r}

```

