---
title: "MEDS Alumni Analysis"
author: "Sofia Sarak"
format: html
---

The following document analyzes the careers of MEDS alumni after graduation. The following explorations were conducted:

- the number of roles an individual has held since graduating (total as well as per year)
- how many of the Class of 2022 are still at their first placement
- average length of job search
- mobility between sectors
- TO DO: time to first placement
- TO DO: % of alumni with missing data
- TO DO: common job titles: analyst, scientist, specialist, planner, engineer, manager
- TO DO: differences in sector, time to first placement, by cohort

```{r}
#| message: false
#| warning: false
#| echo: false

# save Bren blue
blue <- "#14375C"

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      data import and initial wrangling                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# load necessary libraries
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggsankey)

# read in data
data <- read_csv(here("data", "meds_2020-26_placement_data.csv")) %>% 
  clean_names()

# encode "N/A" as actual NA values

data <- data %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(initial_placement_job_title:fourth_placement_end_date, ~na_if(.x, "Unknown")))

```

## Placements Since Graduation

This analysis includes promotions within the same organization.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##        placement data wrangling - create binary placement variable       ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# create placement data frame (for binary vars)
placement <- data %>% 
  
  # keep only necessary columns (a placement job title signifies placement)
  select(year, last_name, first_name, 
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 

# keep only identifiers and binary vars
  
  select(year, last_name, first_name, initial_bin, second_bin, third_bin, fourth_bin)

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                find total number of placements per individual            ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement <- placement %>% 
  
  # sum columns for total positions
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                  find placements per individual per year                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placement <- placement %>% 
  
  mutate(
    
    # calculate number of years since grad (years in workforce)
    workforce_years = 2026 - year,
         
    # divide placements by number of years in workforce
    placements_per_year = case_when(
      
      # for MEDS class of 2026, default to total placement (otherwise, 1/0 = Inf)
      workforce_years == 0 ~ total_placements,
      
      # division for other years
      .default = total_placements/workforce_years))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

placements_plot <- placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(total_placements)) +
    geom_bar(fill = blue,) +
  labs(x = "Number of Placements Since Graduation",
       y = "Number of Graduates",
       title = "Job Mobility in MEDS Graduates") +
  theme_minimal()

yearly_placements_plot <- placement %>% 
  
  # remove individuals with no placements
  filter(total_placements != 0) %>% 
  
  # plot distribution
  ggplot(aes(placements_per_year)) +
    geom_bar(fill = blue) +
  labs(x = "Number of Placements per Year",
       y = "Number of Graduates",
       title = "Job Mobility in MEDS Graduates",
       subtitle = "(Yearly Average)") +
  theme_minimal()

placements_plot + yearly_placements_plot
```
The majority of graduates switch positions once a year or less. This is skewed heavily by graduates who have only had a year or two in the workforce.

## Class of 2022: Have they retained their first placement?

Here, we isolate the Class of 2022 -- they are the first MEDS cohort, and have had 5 years in the workforce. This provides a more standardized look at job mobility over time.

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##            create a placement data frame just for class of 2022          ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# this version of the df filters for 2022 and also keeps organization name variables

first_cohort <- data %>% 
  
  # filter for class of 2022
  filter(year == 2022) %>% 
  
  # select necessary columns
  select(year, last_name, first_name, initial_placement_organization_name, current_most_recent_position_organization_name,
         
         # placement columns to know if there were any promotions
         initial_placement_job_title, second_placement_job_title, 
         third_placement_job_title, fourth_placement_job_title) %>% 
  
  # create binary columns based on NA values SOMETHING WRONG WITH IT THIS TIME?
  mutate(
     # initial placement
    initial_bin = case_when(is.na(initial_placement_job_title) ~ 0,
                                 .default = 1),
    
    # second placement
    second_bin = case_when(is.na(second_placement_job_title) ~ 0,
                                 .default = 1),
    
    # third placement
    third_bin = case_when(is.na(third_placement_job_title) ~ 0,
                                 .default = 1),
    
    # fourth placement
    fourth_bin = case_when(is.na(fourth_placement_job_title) ~ 0,
                                 .default = 1)
    ) %>% 
  
  select(year, last_name, first_name,  initial_placement_organization_name, current_most_recent_position_organization_name,
         initial_bin, second_bin, third_bin, fourth_bin) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                  find individuals at their original orgs                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

original_org <- first_cohort %>% 
  
  # keep only obvs where initial and current org are the same
  filter(initial_placement_organization_name == current_most_recent_position_organization_name) %>% 
  
  # add column for the total number of promotions/position changes
  mutate(total_placements = rowSums(pick(ends_with("bin")), na.rm = TRUE))
```

Insert conclusions

## Length of Job Search

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        add job search length columns                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

search <- data %>% 
  
  # select necessary columns
  select(year, last_name, first_name, ends_with("date")) %>% 
  
  # transform date columns to date format
  mutate(across(ends_with("date"), my)) %>% 
  
  # add column for graduation day approximation (06 + `year` + 15)
  mutate(grad_date = paste0("06/", year, "/15"),
         grad_date = myd(grad_date)) %>% 
  
  # calculate amount of time between graduation and initial placement
  mutate(initial_search_days = as.numeric(initial_placement_start_date - grad_date),
         
         # fix so that if job was started in July (14 days or earlier), search length is actually 0
         initial_search_days = if_else(initial_search_days < 15, 0, initial_search_days))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##             filter for jobs found before/after grad, or at all           ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Select only for positions that were found after graduation (> 0 days), then find the number of observations
after_grad <- search %>% 
  filter(initial_search_days > 0)
after_grad <- nrow(after_grad)

# Select only for positions that were found before graduation (= 0 days), then find the number of observations
before_grad <- search %>% 
  filter(initial_search_days == 0)
before_grad <- nrow(before_grad)

# For a more accurate representation of job search length, we only select for those individuals that have found positions/the positions are publicly available
search_non_na <- search %>% 
  filter(!is.na(initial_search_days))

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                 create viz                               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

search_non_na %>% 
  ggplot(aes(x = initial_search_days)) + 
  geom_histogram(fill = blue) +
  labs(x = "Initial Job Search Length (Days)",
       y = "Number of Graduates",
       title = "Distribution of Job Search Lengths in MEDS Graduates",
       subtitle = "Average job search length of 62 days is shown in red") + 
  theme_minimal() +
  
  # add line for average
  geom_vline(xintercept = mean(search$initial_search_days, na.rm = TRUE),
             color = "red",
             linetype = "dashed")
```
type conclusions here

including the following stats: 

- number of roles found before or less than a month after graduation vs number of roles found after

also, include assumption that positions started in July were found before graduation (job search = 0) due to lack of granularity in data.

## Mobility between sectors

```{r}
#| message: false
#| warning: false
#| echo: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      read in and wrangle sector data                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

sector <- read_csv(here("data", "sectors.csv")) %>% 
  
  clean_names() %>% 
  
  # replace "N/A" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "N/A"))) %>% 

  # replace "Unknown" string with NA value, in placement columns
  mutate(across(everything(), ~na_if(.x, "Unknown"))) %>% 
  
  # remove rows where there was no initial placement/no data (mostly class of 2026)
  filter(!is.na(initial_placement))

# convert data to ggsankey format
sector_sankey <- sector %>% 
  
  # function comes from ggsankey package
  make_long(initial_sector, second_sector, third_sector, fourth_sector) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        plot using ggsankey package                       ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

ggplot(sector_sankey, aes(x = x, # set aesthetics to make make_long() formatting
                         next_x = next_x,
                         node = node,
                         next_node = next_node,
                         fill = factor(node),
                         label = node)) +
         
         # customize 
         geom_sankey(flow.alpha = 0.5,  # opacity
                     node.color = "NA", # border color
                     show.legend = FALSE) + 
         
         # add node labels
         geom_sankey_label(size = 3, color = "black", fill= "white") +
  
        # set x-axis labels
         scale_x_discrete(labels = c("Initial\nPlacement", "Second\nPlacement", "Third\nPlacement", "Fourth\nPlacement")) +
  
        theme_bw()+
        theme(legend.position = "none",
              axis.title = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              panel.grid = element_blank()) +
  
        scale_fill_viridis_d(option = "inferno") +
  
        labs(title = "Movement across sectors",
             subtitle = "MEDS alumni, 2020-2026",
             fill = "Nodes")
```

